---
title: "R Notebook"
output: html_notebook
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


```{r}
# R7-1.R: カイ自乗検定
obtable2 <- matrix(c(25, 23, 12, 15, 17, 8),
                   nrow = 2, byrow = TRUE)
obtable2

chisq.test(obtable2)
```

```{r}
# R7-2.R: ANOVA
# 3群の群別サンプル数、平均、SD
nx <- c(40, 40, 20); mb <- c(50, 60, 45); sdb <- c(50, 50, 50)*0.01

# データ生成
t1 <- rnorm(nx[1], mb[1], sdb[1])
t2 <- rnorm(nx[2], mb[2], sdb[2])
t3 <- rnorm(nx[3], mb[3], sdb[3])

tb <- c(t1, t2, t3)

x <- c(rep("x1", nx[1]), rep("x2", nx[2]), rep("x3", nx[3]))
boxplot(tb ~ x)
```

```{r}
# ANOVA
summary(aov(tb ~ x ))
```

```{r}
# R7-5.R 固有値分解

Nm <- 100 # 項目数
Ns <- c(rpois(4, 100), rpois(20, 10))   # rpois(サンプリング回数, λ)
Npop <- length(Ns)

Ns
Npop   # 集団数
```

```{r}
M <- NULL   # 全ジェノタイプを収める行列

for (j in 1:Npop) {
  tmpM <- matrix(rep(0, Nm*Ns[j]), ncol=Nm)
  
  # 亜集団ごとに平均を振っている
  for (i in 1:Nm) {
    af <- rnorm(1, mean = 0, sd = 1)
    tmpM[, i] <- rnorm(Ns[j], af)
  }
  
  M <- rbind(M, tmpM)
}

tmpM
```

```{r}
# データの標準化
wholemean <- mean(M)
M <- M - wholemean
mu <- apply(M, 2, mean)
M <- t(t(M) - mu)   # Rの演算はブロードキャストされる
```

```{r}
# 特異値分解
# Mは正方行列じゃないので注意
svdout <- svd(M)
M2 <- svdout$u %*% diag(svdout$d)

dim(M)
dim(M2)
```

```{r}
dim(svdout$u)
dim(svdout$v)
length(svdout$d)
```

```{r}
tail(svdout$u)
```


```{r}
par(mfcol=c(1,2))

# 固有値分解前後をimageプロット
image(1:sum(Ns), 1:Nm, M)
image(1:sum(Ns), 1:Nm, M2)
```

```{r}
df1 <- as.data.frame(M); df2 <- as.data.frame(M2)
L <- 1:5; par(mfcol=c(1,1))
plot(df1[, L])
plot(df2[, L])
```

```{r}
vM1 <- apply(M, 2, var)
vM2 <- apply(M2, 2, var)
vM1
vM2
```

```{r}
ylim <- c(min(vM1, vM2), max(vM1, vM2))
plot(vM1, ylim = ylim, type = "b")
par(new=T)
plot(vM2, ylim = ylim, type = "b", col = "red")
```





























